### 2014-07-11 manuscripts/supplement.Rmd
tweaking mansucript

---
layout: review, 11pt
title: "Supplementary material for -- RNeXML: Parsing and Serializing the Next Generation of Phyloinformatic Data in R"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Hilmar Lapp
    affiliation: NESCent
  - name: Scott Chamberlain
    affiliation: SimonFraser
  - name: Rutger Vos
    affiliation: NBC
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: NESCent
    address: | 
      National Evolutionary Synthesis Center, Durham, NC, USA
  - code: SimonFraser
    address: |
      Department of Biology, Simon Fraser University, Burnaby, Canada
  - code: NBC
    address: | 
      Naturalis Biodiversity Center, Leiden, the Netherlands

bibliography: components/sup-references.bib
csl: components/methods-in-ecology-and-evolution.csl
documentclass: components/elsarticle

output: 
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: false
    fig_caption: true

---

```{r supplement-compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE, 
               cache = 1, comment = NA, verbose = TRUE)
basename <- 'supplement' 
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

library("knitcitations")
cleanbib()
cite_options(citation_format = "pandoc", check.entries=FALSE)
```

## SPARQL Queries

```{r}
library(rrdf)
library(phytools)
library(RNeXML)
```

```{r}
nexml <- nexml_read(system.file("examples/primates.xml", package="RNeXML"))
```

Extract the RDF graph from the nexml

```{r}
rdf <- get_rdf(system.file("examples/primates.xml", package="RNeXML"))
saveXML(rdf, "rdf_meta.xml") # rrdf requires a file name, so we must write the XML out first
graph <- load.rdf("rdf_meta.xml")
```

fetch the NCBI URI for the taxon that has rank 'Order', i.e. the root of the primates. The dot operator `.` between clauses implies a join, in this case

```{r}
root <- sparql.rdf(graph, 
    "SELECT ?uri WHERE { 
        ?id <http://rs.tdwg.org/ontology/voc/TaxonConcept#rank> <http://rs.tdwg.org/ontology/voc/TaxonRank#Order> . 
        ?id <http://rs.tdwg.org/ontology/voc/TaxonConcept#toTaxon> ?uri    
    }"               
)
```

Define a function to get the name

```{r}
get_name <- function(id) {
    max <- length(nexml@otus[[1]]@otu)
    for(i in 1:max) {
        if ( nexml@otus[[1]]@otu[[i]]@id == id ) {
            label <- nexml@otus[[1]]@otu[[i]]@label
            label <- gsub(" ","_",label)
            return(label)
        }
    }
}
```

 define a recursive function to build newick

```{r}
recurse <- function(node){

    # fetch the taxonomic rank and id string
    rank_query <- paste0(
        "SELECT ?rank ?id WHERE {
            ?id <http://rs.tdwg.org/ontology/voc/TaxonConcept#toTaxon> <",node,"> .
            ?id <http://rs.tdwg.org/ontology/voc/TaxonConcept#rank> ?rank
          }")
    result <- sparql.rdf(graph, rank_query)

    # get the local ID, strip URI part
    id <- result[2]
    id <- gsub("^.+#", "", id, perl = TRUE)

    # if rank is terminal, return the name
    if (result[1] == "http://rs.tdwg.org/ontology/voc/TaxonRank#Species") {
        return(get_name(id))
    }

    # recurse deeper
    else {
        child_query <- paste0(
            "SELECT ?uri WHERE {
                ?id <http://www.w3.org/2000/01/rdf-schema#subClassOf> <",node,"> .
                ?id <http://rs.tdwg.org/ontology/voc/TaxonConcept#toTaxon> ?uri
            }")
        children <- sparql.rdf(graph, child_query)

        # the newick can be made to contain interior node labels by inserting get_name(id), before the sep="" argument
        return(paste("(", 
                     paste(sapply(children, recurse), 
                           sep = ",", collapse = "," ), 
                     ")",  sep = "", collapse = ""))
    }
}
```

build the tree and visualize it

```{r}
newick <- paste(recurse(root), ";", sep = "", collapse = "")
tree <- read.newick(text = newick)
collapsed <- collapse.singles(tree)
plot(collapsed, type = "cladogram")
```

## Understanding the `nexml` S4 object

The `RNeXML` package provides many convenient functions to add and extract
information from `nexml` objects in the R environment without requiring
the reader to understand the details of the NeXML data structure and
making it less likely that a user will generate invalid NeXML syntax
that could not be read by other parsers. The `nexml` object we have been using
in all of the examples is built on R's S4 mechanism. Advanced users may
sometimes prefer to interact with the data structure more directly using 
R's S4 class mechanism and subsetting methods. Many R users are more familiar
with the S3 class mechanism (such as in the `ape` package phylo objects)
rather than the S4 class mechanism used in phylogenetics packages such as 
`ouch` and `phylobase`. The `phylobase` vignette provides an excellent introduction
to these data structures.  Users already familiar with subsetting lists and other
S3 objects in R are likely familar with the use of the `$` operator, such as
`phy$edge`. S4 objects simply use an `@` operator instead (but cannot be subset
using numeric arguments such as `phy[[1]]` or named arguments such as phy[["edge"]]).

The `nexml` object is an S4 object, as are all of its components (slots).  Its 
hierarchical structure corresponds exactly with the XML tree of a NeXML file, with 
the single exception that both XML attributes and children are represented as slots.  
S4 objects have constructor functions to initialize them.  We create a new `nexml` 
object with the command:

```{r}
nex <- new("nexml")
```

We can see a list of slots contained in this object with

```{r}
slotNames(nex)
```

Some of these slots have already been populated for us, for instance, the schema version and default namespaces:

```{r}
nex@version
nex@namespaces
```

Recognize that `nex@namespaces` serves the same role as `get_namespaces` function, but provides direct access to the slot data.  For instance, with this syntax we could also overwrite the existing namespaces with `nex@namespaces <- NULL`.  Changing the namespace in this way is not advised.

Some slots can contain multiple elements of the same type, such as `trees`, `characters`, and `otus`.  For instance, we see that

```{r}
class(nex@characters)
```

is an object of class `ListOfcharacters`, and is currently empty,

```{r}
length(nex@characters)
```

In order to assign an object to a slot, it must match the class definition
of the slot.  We can create a new element of any given class with the
`new` function,

```{r}
nex@characters <- new("ListOfcharacters", list(new("characters")))
```

and now we have a length-1 list of character matrices,

```{r}
length(nex@characters)
```

and we access the first character matrix using the list notation, `[[1]]`. Here we check the class is a `characters` object.

```{r}
class(nex@characters[[1]])
```

Direct subsetting has two primary use cases: (a) useful in looking up (and possibly editing) a specific value of an element, or (b) when adding metadata annotations to specific elements. Consider the example file

```{r}
f <- system.file("examples", "trees.xml", package="RNeXML")
nex <- nexml_read(f)
```

We can look up the species label of the first `otu` in the first `otus` block:

```{r}
nex@otus[[1]]@otu[[1]]@label
```

We can add metadata to this particular OTU using this subsetting format

```{r}
nex@otus[[1]]@otu[[1]]@meta <- 
  c(meta("skos:note", 
          "This species was incorrectly identified"),
         nex@otus[[1]]@otu[[1]]@meta)
```

Here we use the `c` operator to append this element to any existing meta annotations to this otu.

# Implementation, architecture and quality control

The `RNeXML` package is built for the NeXML Schema, [version 2009],
using the R software environment and the XML package for parsing and
generating XML `r citep(citation("XML"))`,  provides conversions into
the popular `phylo` format provided by the `ape` package `r citep(citation("ape"))`
`RNeXML` can automatically check taxonomic names and add taxonomic
identifiers when generating NeXML using methods from the `taxize` package `r citep(citation("taxize"))`.
Unique identifiers are provided with the `uuid` package `r citep(citation("uuid"))`,
and internet interfaces by the `httr` package `r citep(citation("httr"))`. The
packages `plyr` and `reshape2``r citep(c(citation("plyr"), citation("reshape2")))` are also imported for internal methods .

<!--  Find more consise way to make sure dependencies are cited. 
      How about citing packages that are only on the suggests list? --> 
<!-- Possibly cite all dependencies and suggests from the DESCRIPTION --> 
<!-- Reference any installation and configuration instructions necessary (e.g. API credentials for figshare) -->

The `RNeXML` package performs several levels of validation on NeXML files
written from or read into R: validating the XML, and validating against
the current NeXML schema, and performing additional checks through the 
NeXML online validation tool for elements that cannot be enforced by
schema validation alone.  Validation serves to improve consistency in
data encoding which facilitates the use of NeXML across a wide variety
of platforms and software.  A lack of a consistent standard and validation
mechanism has continually plagued the current NEXUS standard and made 
interoperability difficult `r citep("Vos NeXML 2012")`.

Several mechanisms are in place to evaluate the quality of the `RNeXML`
package functions themselves.  In addition to the examples from the
documentation and the automated package checks provided for the R system,
the `RNeXML` package includes a complete test suite using the `testthat`
package `r citep(citation("testthat"))`. Package development on Github includes continuous 
integration through the `travis` platform, in which the full test suite
is run every time the code is changed.  This not only gives users confidence
that the development copy can be successfully installed, but should assist
other developers seeking to contribute to the package to test that their proposed 
changes have not negatively impacted any of the existing code base.

```{r cleanup, include=FALSE, cache=FALSE}
write.bibtex(file="components/sup-references.bib")
unlink("simmap.xml")
unlink("birds.xml")
unlink("geospiza.xml")
unlink("rdf_meta.xml")
```

<!-- MISC / SCRATCH

What to say about existing formats?  
A summary of some phylogenetic formats in R:

| Class     | R package   | R Object 
|-----------|-------------|-------
| phylo     | ape         | S3 
| phylo4    | phylobase   | S4
| phylo4d   | adephylo    | S4
| ouchtree  | ouch        | S4 
| treeshape | apTreeshape | list

See [phylobase vignette, 2010]

phylobase::phylo4 was an attempt to solve the problems associated with
the popular ape::phylo format (unpredictable errors, loose and multiple
conflicting specifications).

-->