### 2015-01-05 manuscripts/manuscript.Rmd
Tightening up various places in package description.

---
#layout: preprint
layout: review, 11pt
title: "RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Hilmar Lapp
    affiliation: NESCent
  - name: Scott Chamberlain
    affiliation: ropensci
  - name: Rutger Vos
    affiliation: NBC
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: NESCent
    address: | 
      National Evolutionary Synthesis Center, Durham, NC, USA
  - code: ropensci
    address: |
      University of California, Berkeley, CA, USA
  - code: NBC
    address: | 
      Naturalis Biodiversity Center, Leiden, the Netherlands
abstract: | 
      1. NeXML is a new exchange format for phylogenetic data and
      metadata that offers substantial improvements in reliability,
      extensibility, and richness over earlier formats such as
      NEXUS. In this paper we present the RNeXML package which
      provides a bridge between this data format and the extensive
      library of phylogenetic tools already available in the popular
      data analysis environment and programming language R, with a
      particular emphasis on comparative methods.

      2. The ability to parse and write NeXML from within R is only a
      small part of the functionality offered by the package.  We
      illustrate how RNeXML facilitates adding and manipulating data
      annotations whose meaning can be defined in a precise and
      machine-interpretable way by means of formal ontologies and other
      semantic web technologies.

      3. An essential feature of NeXML is to provide both the
      flexibility to be continually extended to meet the emerging
      needs of the phylogenetics community without breaking software
      unaware of such extensions. We illustrate how RNeXML makes
      it easy to extend NeXML in this manner.

      4. RNeXML is well suited for users to preserve and maintain
      their personal archives of phylogenetic and comparative trait
      data in consistent, query-able, and cross-platform format, to
      leverage the rapidly growing public community archives of rich
      NeXML files, and to interact with web services that return data
      and their metadata packaged together as NeXML.

bibliography: components/references.bib
csl: components/methods-in-ecology-and-evolution.csl
documentclass: components/elsarticle

output: 
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true

---

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE, 
               cache = 1, comment = NA, verbose = TRUE)
basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile')) 
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

```

# Introduction

Users of the popular statistical and mathematical computing platform R
[@R] enjoy a wealth of readily installable comparative phylogenetic
methods and tools <!-- needs a citation to support it -->. Exploiting
the opportunities arising from this wealth for complex and integrative
comparative research questions relies on the ability to reuse and
integrate previously generated or published data and metadata. The
expanding data exchange needs of the evolutionary research
community are rapidly outpacing the capabilities of most current
and widely used data exchange standards [@Vos_2012], which were all
developed a decade or more ago. This has resulted in a radiation of
different data representations and exchange standard "flavors" that
are no longer interoperable at the very time when the growth of
available data and methods has made that interoperability most
valuable. In response to the unmet needs for standardized data
exchange in phylogenetics, a modern XML-based exchange standard,
called NeXML, has recently been developed [@Vos_2012]. NeXML
comprehensively supports current data exchange needs, is predictably
machine-readable, and is forward compatible.

The exchange problem for phylogenetic data is particularly acute in
light of the challenges in finding and sharing phylogenetic data
without the otherwise common loss of most data and metadata semantics
[@Drew_2013; @Stoltzfus_2012; @Cranston_2014]. For example, the still popular
NEXUS file format [@Maddison_1997] cannot consistently represent
horizontal gene transfer or ambiguity in reading a character (such as
a DNA sequence base pair).  This and other limitations have led to modifications of NEXUS in
different ways for different needs, with the unfortunate result that
NEXUS files generated by one program can be incompatible with another
[@Vos_2012]. Without a formal grammar, software based on NEXUS files
may also make inconsistent assumptions about tokens, quoting, or
element lengths.  @Vos_2012 estimates that as many as 15% of the NEXUS
files in the CIPRES portal contain unrecoverable but hard to diagnose
errors.

A detailed account of how the NeXML standard addresses these and other
relevant challenges can be found in @Vos_2012. In brief, NeXML was
designed with the following important properties. First, NeXML is
defined by a precise grammar that can be programmatically
**validated**. This ensures that any NeXML file can be expected to be
parsed and read without errors by any NeXML-compliant software,
provided the file passes validation. Second, NeXML is **extensible**:
a user can define representations of new, previously unanticipated
information (as we will illustrate) without violating its defining
grammar. Third and most importantly, NeXML is **rich in computable
semantics**. The formal semantics used in NeXML is machine-readable in
a predictable manner, and also enables machine reasoning about the
data and metadata (see @Parr2011), especially when metadata attributes
and values are anchored in appropriately axiomatized ontologies.

`RNeXML` not only allows R's ecosystem to read and write data in
the NeXML format through an interface that is no more involved than
reading or writing data from other phylogenetic data formats, but also
carries immediate benefits for its users. For example, comparative
analysis R packages and users frequently add their own metadata
annotations to the phylogenies they work with, such as annotations
of species, stochastic character maps, trait values, model estimates
and parameter values. `RNeXML` affords R the capability to harness
machine-readable semantics and an extensible metadata schema to
capture, preserve, and share these and other kinds of information, all
through an API instead of having to understand in detail the schema
underlying the NeXML standard. To assist users in meeting the rising
bar for best practices in data sharing in phylogenetic research
[@Cranston_2014], `RNeXML` captures metadata information from the R
environment to the extent possible, and applies reasonable
defaults.

# The RNeXML package

The `RNeXML` package is written entirely in R and available under a
Creative Commons Zero public domain waiver. The current development version can be found
on Github at [https://github.com/ropensci/RNeXML](), and the stable
version can be installed from the CRAN repository.

```{r echo=FALSE}
install.packages("RNeXML", dependencies=TRUE)
library(RNeXML)
```

`RNeXML` is part of the rOpenSci project. Users of `RNeXML` are
encouraged to submit bug reports or feature requests in the issues log
on Github, or to contact `ropensci-discuss@googlegroups.com` or the
phylogenetics R users group list at `r-sig-phylo@r-project.org` for
help.

## Representation of NeXML documents in R

Conceptually, a NeXML document has the following components: (1)
phylogeny topology data, (2) character or trait data in matrix form, (3)
operational taxonomic units (OTUs), and (4) metadata. To represent the
contents of a NeXML document (currently in memory), `RNeXML` defines
the `nexml` object type. This type therefore holds phylogenetic trees
as well as character or trait matrices, and all metadata, which is
similar to the phylogenetic data object types defined in the
`phylobase` package [@phylobase], but contrasts with the more widely
used ones defined in the
`ape` package [@Paradis_2004], which represents trees alone.

When reading and writing NeXML documents, `RNeXML` aims to map their
components to and from, respectively, their most widely used
representations in R. As a result, the types of objects accepted or returned
by the package's methods are the `phylo` and `multiPhylo` objects from
the `ape` package [@Paradis_2004] for phylogenies, and R's native
`data.frame` list structure for data matrices.

## Reading phylogenies and character data

The method `nexml_read()` reads NeXML files, either from a local
file, or from a remote location via its URL, and returns an object of
type `nexml`:

```{r}
nex <- nexml_read("components/trees.xml")
```

The method `get_trees_list()` can be used to extract the phylogenies
as an `ape::multiPhylo`
object, which can be treated as a list of `ape::phylo` objects:

```{r}
phy <- get_trees_list(nex)
```

The `get_trees_list()` method is designed for use in scripts, providing
a consistent and predictable return type regardless of the number of
phylogenies a NeXML document contains. For greater convenience in
interactive use, the method `get_trees()` returns the R object most
intuitive given the arrangement of phylogeny data in the source NeXML
document. For example, the method returns an `ape::phylo` object if the NeXML
document contains a single phylogeny, an `ape::multiPhylo` object if
it contains multiple phylogenies arranged in a single `trees` block,
and a list of `ape::multiPhylo` objects if it contains multiple
`trees` blocks (the capability for which NeXML inherits from NEXUS).

If the location parameter with which the `nexml_read()` method is
invoked is recognized as a URL, the method will automatically download the
document to the local working directory and read it from there. This
gives convenient and rapid access to phylogenetic data published in
NeXML format on the web, such as the content of the
phylogenetic data repository TreeBASE [@Piel_2002]. For
example, the following plots a tree in TreeBASE (using ape's plot function):

```{r eval=FALSE fig.show="hide"}
tb_nex <- nexml_read(
"https://raw.github.com/TreeBASE/supertreebase/master/data/treebase/S100.xml")
tb_phy <- get_trees_list(tb_nex)
plot(tb_phy[[1]]) 
```

The method `get_characters()` obtains character data matrices from a
`nexml` object, and returns them as a standard `data.frame` R object
with columns as characters and rows as taxa:

```{r }
nex <- nexml_read("components/comp_analysis.xml")
get_characters(nex)
```

A NeXML data matrix can be of molecular (for molecular sequence
alignments), discrete (for most morphological character data), or
continuous type (for many trait data). To enable strict validation of data types NeXML allows, and if their
data types differ requires multiple data matrices to be separated into
different "blocks". Since
the `data.frame` data structure in R has no such constraints, the
`get_characters()` method combines such blocks as separate columns
into a single `data.frame` object, provided they correspond to the
same taxa. Otherwise, a list of `data.frame`s is returned, with list
elements corresponding to characters blocks. Similar to the methods
for obtaining trees, there is also a method `get_characters_list()`,
which always returns a list of `data.frame`s, one for each character
block.

## Writing phylogenies and character data

The method `nexml_write()` generates a NeXML file from its input parameters. In
its simplest invocation, the method writes a tree to a file:

```{r}
data(bird.orders)
nexml_write(bird.orders, file = "birds.xml")
```

The first argument to `nexml_write()` is either an object of type `nexml`, or any
object that can be coerced to it, such as in the above example an
`ape::phylo` phylogeny. Alternatively, passing a `multiPhylo` object
would write a list of phylogenies to the file.

In addition to trees, the `nexml_write()` method also allows to specify
character data as another parameter. The following example uses data
from the comparative phylogenetics R package `geiger` [@Harmon_2008].

```{r}
library("geiger")
data(geospiza)
nexml_write(trees = geospiza$phy, 
            characters = geospiza$dat,
            file="geospiza.xml")
```

## Creating and populating `nexml` objects

Instead of packaging the various components for a NeXML file at the
time of writing the file, `RNeXML` also allows users to create and
iteratively populate in-memory `nexml` objects. The methods to do
this are `add_characters()`, `add_trees()`, and `add_meta()`, for
adding characters, trees, and metadata, respectively. Each of these
functions will automatically create a new nexml object if not supplied
with an existing one as the last (optional) argument.

For example, here we use `add_trees()` to first create a `nexml`
object with the phylogeny data, and then add the character data to it:

```{r}
nexObj <- add_trees(geospiza$phy)
nexObj <- add_characters(geospiza$dat, nexObj)
```

The data with which a `nexml` object is populated need not share the
same OTUs.  `RNeXML` automatically adds new, separate OTU blocks into
the NeXML file for each data matrix and tree that uses a different set of OTUs.

Other than storage size, there is no limit to the number of
phylogenies and character matrices that can be included in a
single NeXML document. This allows, for example, to capture samples from a
posterior probability distribution of inferred or simulated phylogenies and
character states in a single NeXML file.

## Data documentation and annotation with built-in metadata

NeXML allows attaching ("_annotating_") metadata to any data element,
and even to metadata themselves. Whether at the level of the document
as a whole or an individual data matrix or phylogeny, metadata can
provide bibliographic and provenance information, for example about
the study as part of which the phylogeny was generated or applied,
which data matrix and which methods were used to generate it. Metadata
can also be attached to very specific elements of the data, such as
specific traits, individual OTUs, nodes, or even edges of the
phylogeny.

As described in @Vos_2012,  to encode metadata annotations NeXML uses
the "Resource Description Framework in Annotations" (RDFa)
[@W3C_2014]. This standard provides for a strict machine-readable
format yet enables future backwards compatibility with
compliant NeXML parsers (and thus `RNeXML`), because the capacity of a
tool to _parse_ annotations is not predicated on _understanding_
the meaning of annotations it has not seen before.

To lower the barriers to sharing well-documented phylogenetic data,
`RNeXML` aims to make recording useful and machine-readable metadata
easier at several levels.

First, when writing a NeXML file the package adds certain basic
metadata automatically if they are absent, using default
values consistent with recommended best practices
[@Cranston_2014]. Currently, this includes naming the software
generating the NeXML, a time-stamp of when a tree was produced, and an
open data license.

Second, `RNeXML` provides a simple method, called
`add_basic_metadata()`, to set metadata attributes commonly
recommended for inclusion with data to be publicly archived or shared
[@Cranston_2014]. The currently accepted parameters include `title`,
`description`, `creator`, `pubdate`, `rights`, `publisher`, and
`citation`. Behind the scenes the method automatically anchors these
attributes in common vocabularies (such as Dublin Core).

Third, `RNeXML` integrates with the R package `taxize`
[@Chamberlain_2013] to mitigate one of the most common obstacles to
reuse of phylogenetic data, namely the misspellings and inconsistent
taxonomic naming with which OTU labels are often fraught. The
`taxize_nexml()` method in `RNeXML` uses `taxize` to match OTU labels
against the NCBI database, and, where a unique match is found, it
annotates the respective OTU with the matching NCBI identifier.

## Data annotation with custom metadata

The RDFa standard adopted by NeXML (see @Vos_2012) allows for
arbitrary custom extensions to meet ongoing scientific needs without
requiring modifications (or "extensions") of the syntax. Specifically,
every metadata annotation consists of a metadata property and a value
of the property; the bearer of the property is determined by where the
annotation appears in the NeXML document (technically, the containing
element in the XML document model). This structure corresponds
directly to Resource Description Framework (RDF)
subject-predicate-object triples, and therefore the only requirement
for custom metadata is that they be cast as RDF triples.

The `RNeXML` interface described above for built-in metadata annotations
perform this cast automatically behind the scenes, including mapping
metadata attributes to terms in requisite common vocabularies (such as
Dublin Core for "title", "creator", etc.) or ontologies. Giving
metadata properties (and where applicable, values) as vocabulary or ontology
terms rather than simple text strings is crucial for allowing machines
to not only parse but also interpret and potentially reason over their
semantics. To achieve this benefit for custom metadata extensions, the
user necessarily needs to handle certain technical details from which
the `RNeXML` API shields her otherwise, in particular the globally unique
identifiers (normally HTTP URIs) of metadata terms and vocabularies. To be consistent with XML
terminology, `RNeXML` calls vocabulary URIs
_namespaces_, and their abbreviations _prefixes_. For example, the
namespace for the Dublin Core Metadata Terms vocabulary is "http://purl.org/dc/elements/1.1/". Using
its common abbreviation "dc", a metadata property "dc:title" expands to the
identifier "http://purl.org/dc/elements/1.1/title". This URI resolves
to a human and machine-readable (depending on access) definition of precisely
what the term `title` in Dublin Core means. In contrast, just using
the text string "title" could also mean the title of a person, a legal
title, the verb title, etc.  URI identifiers of metadata vocabularies
and terms are not mandated to resolve, but if machines are to derive
the maximum benefit from them, they should resolve to a definition of
their semantics in RDF.

`RNeXML` includes methods to obtain and manipulate metadata
properties, values, identifiers, and namespaces. The
`get_namespaces()` method accepts a `nexml` object and returns a named list
of namespace prefixes and their corresponding identifiers known to the object:

```{r}
prefixes <- get_namespaces(birds)
prefixes["dc"]
```

The `get_metadata()` method returns, as a named list, the metadata
annotations for a given `nexml` object at a given level, with the
whole NeXML document being the default level (`"all"` extracts all metadata objects):

```{r}
meta <- get_metadata(birds) 
otu_meta <- get_metadata(birds, level="otu")
```

The returned list does not include the data elements to which the
metadata are attached. Therefore, a different approach, documented in
the package vignette, is recommended for accessing the metadata
attached to data elements.

The `meta()` method creates a new metadata object from a poperty name
and content (value). For example, the following creates a modification
date metadata object, using a property in the PRISM vocabulary:

```{r}
modified <- meta(property = "prism:modificationDate", content = "2013-10-04")
```

Metadata annotatuons in `NeXML` can be nested within another
annotation, which the `meta()` method accommodates by accepting a
parameter `children`, with the list of nested metadata objects (which
can themselves be nested) as value.

The `add_meta()` function adds metadata objects as annotations to a
`nexml` object at a specified level, with the default level being the
NeXML document as a whole:

```{r}
birds <- add_meta(modified, birds) 
```

If the prefix used by the metadata property is not among the built-in
ones (which can be obtained using `get_namespaces()`), it has to be provided along with its URI as the `namespaces`
parameter. For example, the following uses the SKOS vocabulary (which
currently is not built-in) to add a note to the trees in the `nexml`
object:

```{r}
history <- meta(property = "skos:historyNote",
  content = "Mapped from the bird.orders data in the ape package using RNeXML")
birds <- add_meta(history, 
                birds, 
                level = "trees",
                namespaces = c(skos = "http://www.w3.org/2004/02/skos/core#"))
```

Alternatively, additional namespaces can also be added in batch using
the `add_namespaces()` method.

By virtue of subsetting the S4 `nexml` object, `RNeXML` also offers
fine control of where a `meta` element is added, for which the package
vignette contains examples.

Because NeXML expresses all metadata using the RDF standard, and
stores them compliant with RDFa, they can be extracted as an RDF
graph, queried, analyzed, and mashed up with other RDF data, local or
on the web, using a wealth of off-the-shelf tools for working with RDF
(see @W3C_2014 or @Hartig_2012). Examples for these possibilities are
included in the `RNeXML` vignette, and the package also comes with a
demonstration that can be run from R using the following command:
`demo("sparql", "RNeXML")`).

## Using metadata to extend the NeXML standard

NeXML was designed to prevent the need for future non-interoperable
"flavors" of the standard in response to new research directions. Its
solution to this inevitable problem is a highly flexible metadata
system without sacrificing strict validation of syntax and structure.

Here we illustrate how `RNeXML`'s interface to NeXML's metadata system
can be used to record and share a type of phylogenetic data not taken
into account when NeXML was designed, in this case stochastic
character maps [@Huelsenbeck_2003]. Such data assign certain parts
(corresponding to time) of each branch in a time-calibrated phylogeny
to a particular "state" (typically of a morphological
characteristic). The current de-facto format for sharing stochastic
character maps, created by `simmap` [@Bollback_2006], a widely used tool
for creating such maps, is a non-interoperable modification of the standard
Newick tree format.

To express these data as a custom metadata
extension, we create metadata annotations for the `edge` elements in
the phylogeny topology. The metadata need to describe the character state
being assigned, and the duration (in terms of branch length) that the edge 
spends in that state. The `NeXML` standard already defines `state`
elements for discrete character traits, which, as most elements in
NeXML, also have unique identifiers by which they can
be referenced. In the following
proposal, we term each series of character state assignments for an edge a
_reconstruction_, and for each state assignment comprising a
reconstruction, we provide the length, the identifier of the character
state, and an ordinal property to record the chronological order of
the assignment (the identifier values in the proposal, "cr1",
"s1", and "s2", are identifiers local to a  -- fictitious -- NeXML document):

```{r}
 m <- meta("simmap:reconstructions", children = c(
        meta("simmap:reconstruction", children = c(

          meta("simmap:char", "cr1"),
          meta("simmap:stateChange", children = c(
            meta("simmap:order", 1),
            meta("simmap:length", "0.2030"),
            meta("simmap:state", "s2"))),

          meta("simmap:char", "cr1"),
          meta("simmap:stateChange", children = c(
            meta("simmap:order", 2),
            meta("simmap:length", "0.0022"),
            meta("simmap:state", "s1")))
          ))))
```

The example uses a prefix, `simmap`, to group the newly introduced
metadata properties in a vocabulary, for which the `add_namespace()`
method can be used to give a URI as an identifier:

```{r}
nex <- add_namespaces(c(simmap = "https://github.com/ropensci/RNeXML/tree/master/inst/simmap.md"))
```

Here the URI does not resolve to a fully
machine-readable definition of the terms and their semantics, but it can
nonetheless be used to provide at least a human-readable informal definition
of the terms.

Because the above representation does not modify the NeXML format, it
should be readable -- even if not necessarily interpretable as a
stochastic character mapping -- by any compliant NeXML parser. In
addition, it includes a direct and machine-readable link to metadata
documentation.

Writing out metadata annotations as in the above proposal may seem
tedious. However, by virtue of `RNeXML`'s API it is straightforward to
define functions that encapsulate these details. As a demonstration,
`RNeXML` comes with functions `simmap_to_nexml()` and
`nexml_to_simmap()` that map bidirectionally between NeXML files with
`simmap` metadata according to the above proposal, and the extension
of the `ape::phylo` class devised by @Revell_2012 in the R package
`phytools`. The `RNeXML` package vignette illustrates how this allows
seamlessly generating and reading NeXML files containing `simmap`
data, and using the various functions from `phytools` designed for
`simmap` objects [@Revell_2012].

## Publishing NeXML files from R

Data archiving is increasingly required by scientific journals,
including in evolutionary biology, ecology, and biodiversity
(e.g. @Rausher_2010). The effort involved with preparing and
submitting properly annotated data to archives remains a notable
barrier to the broad adoption of data archiving and sharing as a
normal part of the scholarly publication workflow
[@Tenopir_2011; @Stodden_2014]. In particular, the majority of
phylogenetic trees published in the scholarly record are inaccessible
or lost to the research community [@Drew_2013].

One of `RNeXML`'s aims is to promote the archival of well-documented
phylogenetic data in scientific data repositories, in the form of
NeXML files. To this end, the method `nexml_publish()` provides an API
directly from within R that allows data archival to become a step programmed
into data management scripts. Initially, the method supports the data repository Figshare (http://figshare.com):

```{r eval = FALSE}
doi <- nexml_publish(birds, visibility = "public", repository="figshare")
```

Figshare also supports "private" visibility, allowing to securely backup
data to a repository and to share them with collaborators prior to public release.

# Future Directions

The goals for continued development of `RNeXML` revolve primarily
around three directions. One is better interoperability with other
existing phylogenetic data representations in R, such those found in
the `phylobase` package [@phylobase]. Another is better integration of
the rich metadata semantics found in ontologies defined in the Web
Ontology Language (OWL), including programmatic access to machine
reasoning with such metadata. Finally, and similar to many other R
packages, including those for comparative phylogenetics, in light of
the increasing size of phylogenomic and other datasets mechanisms may
eventually be needed that avoid the requirement to load an entire
file's content into memory.

## Acknowledgements

This project was supported in part by the National Evolutionary
Synthesis Center (NESCent) (NSF #EF-0905606), and grants from the
Natioal Science Foundation (DBI-1306697) and the Alfred P Sloan
Foundation (Grant 2013-6-22). `RNeXML` started as a project idea for
the Google Summer of Code, and we thank Kseniia Shumelchyk for taking
the first steps to implement it. We are grateful to F. Michonneau for
helpful comments on an earlier version of this manuscript.

```{r cleanup, include=FALSE, cache=FALSE}
unlink("simmap.xml")
unlink("birds.xml")
unlink("geospiza.xml")
unlink("rdf_meta.xml")
```

<!-- Footnotes -->

[^1]: By providing the optional `dependencies` argument, R will install
additional packages used in some of the advanced examples shown later
but which are not required for basic functionality.

[^2]: For convenience `nexml_read` is also aliased as `read.nexml`,
for users familiar with the somewhat analogous `read.nexus` function.