### 2018-12-08 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

[![DOI](https://zenodo.org/badge/11856817.svg)](https://zenodo.org/badge/latestdoi/11856817)
[![Build Status](https://api.travis-ci.org/ropensci/RNeXML.png)](https://travis-ci.org/ropensci/RNeXML)
[![Windows build status](https://ci.appveyor.com/api/projects/status/dhiwp5blx2ns2yba/branch/master?svg=true)](https://ci.appveyor.com/project/cboettig/rnexml/branch/master)
[![CRAN status](https://www.r-pkg.org/badges/version/RNeXML)](https://cran.r-project.org/package=RNeXML)
[![codecov.io](https://codecov.io/github/ropensci/RNeXML/coverage.svg?branch=master)](https://codecov.io/github/ropensci/RNeXML?branch=master)
![downloads](http://cranlogs.r-pkg.org/badges/grand-total/RNeXML)

RNeXML: The next-generation phylogenetics format comes to R
=====================================================

* Maintainer: Carl Boettiger
* Authors: Carl Boettiger, Scott Chamberlain, Hilmar Lapp, Kseniia Shumelchyk, Rutger Vos
* License: BSD-3 
* [Issues](https://github.com/ropensci/RNeXML/issues): Bug reports, feature requests, and development discussion.

An extensive and rapidly growing collection of richly annotated phylogenetics data is now available in the NeXML format. NeXML relies on state-of-the-art data exchange technology to provide a format that can be both validated and extended, providing a data quality assurance and adaptability to the future that is lacking in other formats. See [Vos et al 2012](http://doi.org/10.1093/sysbio/sys025 "NeXML: Rich, Extensible, and Verifiable Representation of Comparative Data and Metadata.") for further details on the NeXML format.

How to cite
-----------

RNeXML has been published in the following article:

> Boettiger C, Chamberlain S, Vos R and Lapp H (2016). “RNeXML: A Package for
> Reading and Writing Richly Annotated Phylogenetic, Character, and Trait Data
> in R.” _Methods in Ecology and Evolution_, **7**, pp. 352-357.
> [doi:10.1111/2041-210X.12469](http://doi.org/10.1111/2041-210X.12469)

Although the published version of the article is paywalled, the source of the manuscript, and a much better rendered PDF, are included in this package (in the `manuscripts` folder). You can also find it [freely available on arXiv](http://arxiv.org/abs/1506.02722).

```{r compile-settings, include=FALSE}
library(methods)
```

Installation
---------------

The latest stable release of RNeXML is on CRAN, and can be installed with the usual `install.packages("RNeXML")` command.  Some of the more specialized functionality described in the Vignettes (such as RDF manipulation) requires additional packages which can be installed using:

```{r eval=FALSE}
install.packages("RNeXML", deps = TRUE)
```

The development version can be installed using:

```{r eval=FALSE}
remotes::install_github("ropensci/RNeXML")
```

Getting Started
----------------

See the vignettes below for both a general quick start and an overview of more specialized features.

- [A Brief Introduction to RNeXML](https://ropensci.github.com/RNeXML/articles/intro)
- [RNeXML: A Package for Reading and Writing Richly Annotated Phylogenetic, Character, and Trait Data in R](https://ropensci.github.com/RNeXML/articles/manuscript.pdf) (published in MEE).
- [Handling Metadata in RNeXML](https://ropensci.github.com/RNeXML/articles/metadata)
- [The `nexml` S4 Object](https://ropensci.github.com/RNeXML/articles/S4)
- [Semantic data & SPARQL with RNeXML](https://ropensci.github.com/RNeXML/articles/SPARQL)
- [Extending NeXML: an example based on simmap](https://ropensci.github.com/RNeXML/articles/simmap)

-----

[![ropensci footer](http://ropensci.org/public_images/github_footer.png)](http://ropensci.org)